{\bf Stateless Model Checking \\ with Data-Race Preemption Points}

### 2016-03-01 oopsla/calcm.tex
write more for calcm

\begin{frame}{MC State of the Art}
	Modern stateless MC tools can {\em mitigate} exponential explosion.
	\begin{itemize}
		\item Dynamic Partial Order Reduction\related{Flanagan '05}
		\item State space estimation\related{Simsa '12}
	\end{itemize}
	\linegap

@@

			& \\
			& \small \texttt{lock(mx);} \\
			& \small \texttt{unlock(mx);} \\
			& \small \texttt{x++;} // B \\
			& \\
			\hline
		\end{tabular}
	\end{center}
	Pure HB: \texttt{mx} establishes $HB(A, B)$; no data race reported (false negative).
	\linegap

	Limited HB: $\neg HB(A, B)$; data race is reported (correctly).
\end{frame}

@@

			& \small \texttt{lock(mx);} \\
			& \small \texttt{bool z = y;} \\
			& \small \texttt{unlock(mx);} \\
			& \small \texttt{if (z);} \\
			& \small \texttt{~~x++;} // B \\
			\hline
		\end{tabular}
	\end{center}
	Pure HB: \texttt{mx} establishes $HB(A, B)$; no data race reported (correctly).
	\linegap

	Limited HB: $\neg HB(A, B)$; data race is reported (false positive).
\end{frame}

@@

\begin{frame}{Iterative Deepening}
	If time allows, combine PPs into larger, more comprehensive state spaces.
	\begin{center}
		\includegraphics[width=0.8\textwidth]{tree3.pdf}
	\end{center}
	% say: and depending on your time budget, you might decide that you expect this one to be feasible, or you might fall back to a smaller state space from before.
\end{frame}

\begin{frame}{Total Verification}
	% Let's come back to this issue of ...
	MC can {\em totally verify} a test by testing a state space with all possible PPs.
	\begin{itemize}
		\item 
	\end{itemize}
	%One downside of hard-coding PPs in advance is timing out when the test is too large, but another downside is if the test is small enough, you might miss a possible full verification!
\end{frame}

@@

>>>>

\begin{frame}{MC State of the Art}
	Modern stateless MC tools can {\em mitigate} exponential explosion.
	\begin{itemize}
		\item Dynamic Partial Order Reduction (DPOR)\related{Flanagan '05}
		\item State space estimation\related{Simsa '12}
	\end{itemize}
	\linegap

@@

			& \\
			& \small \texttt{lock(mx);} \\
			& \small \texttt{unlock(mx);} \\
			& \small \texttt{x++;} // B \\
			& \\
			\hline
		\end{tabular}
	\end{center}
	{\bf Pure HB}: \texttt{mx} establishes $HB(A, B)$; no race reported (false negative).
	\linegap

	{\bf Limited HB}: $\neg HB(A, B)$; race is reported (correctly).
\end{frame}

@@

			& \small \texttt{lock(mx);} \\
			& \small \texttt{bool z = y;} \\
			& \small \texttt{unlock(mx);} \\
			& \small \texttt{if (z);} \\
			& \small \texttt{~~x++;} // B \\
			\hline
		\end{tabular}
	\end{center}
	{\bf Pure HB}: \texttt{mx} establishes $HB(A, B)$; no race reported (correctly).
	\linegap

	{\bf Limited HB}: $\neg HB(A, B)$; race is reported (false positive).
\end{frame}

@@

\begin{frame}{Iterative Deepening}
	If time allows, combine PPs into larger, more comprehensive state spaces, eventually converging to using all PPs at once ({\em ``maximal'' state space}).
	\begin{center}
		\includegraphics[width=0.8\textwidth]{tree3.pdf}
	\end{center}
	% say: and depending on your time budget, you might decide that you expect this one to be feasible, or you might fall back to a smaller state space from before.
\end{frame}

\begin{frame}{Total Verification}
	% Let's come back to this issue of ...
	MC can {\em totally verify} a test by testing a state space with all possible PPs.
	\begin{itemize}
		\item Ordinarily infeasible, but some tests are small enough
		\item Whether a test can be totally verified can't be known in advance.
		\item ``Hard-coded'' PP approach\related{Musuvathi '08} sacrifices this possibility.
	%One downside of hard-coding PPs in advance is timing out when the test is too large, but another downside is if the test is small enough, you might miss a possible full verification!
		\item What PP strategy works for both small and large tests?
	\end{itemize}
	\pause
	\linegap

	Synchronization API PPs \& data-race PPs are sufficient!
	\begin{itemize}
		\item {\bf Theorem (Convergence)}:
			{\em If a bug can be exposed by any thread interleaving
				possible by preempting on all instructions,
				% during a specific test
			Quicksand will eventually test an equivalent interleaving which exposes the same bug.}
		\item Proof appeals to DPOR soundness\related{Flanagan '05} and Limited HB.
	\end{itemize}
\end{frame}

@@

\breakslide{\Huge Evaluation}

\begin{frame}{Evaluation}
	Evaluation questions:
	\begin{enumerate}
		\item Do we find more bugs than prior MC with data-race PPs?
		\item Do we find bugs faster than prior MC, even without data-race PPs?
			% say: "the same bugs"
		\item Does MC uncover new, nondeterministic, data-race candidates?
		\item Do we provide more total verifications than prior MC?
	\end{enumerate}
\end{frame}

\begin{frame}{Experimental Setup}
	\textbf{Test suite}: 79 ``P2''s $\times$ 6 test cases + 78 ``Pintos''s $\times$ 3 test cases
	\begin{itemize}
		\item P2: CMU 15-410 thread library project (\textasciitilde{}1800 LOC)
		\item Pintos: Berkeley CS162/U. Chicago CS320 kernel project (\textasciitilde{}700 LOC)
		\item 629 unique state spaces (i.e., project/test pairs)
	\end{itemize}
	\pause
	\linegap

	{\bf Quicksand experiment}: 1 hour $\times$ 10 CPUs per test
	\begin{itemize}
		\item {\bf QS}: Data-race PPs enabled
		\item {\bf QS-no-DRs}: Restrict iterative deepening to sync API PPs only
	\end{itemize}
	\linegap

	{\bf Control experiment}: 10 hours $\times$ 1 CPU per test
	\begin{itemize}
		\item Test only ``maximal`` state space, using all sync API PPs at once
		\item {\bf DPOR}: Use DPOR\related{Flanagan '05} to test maximal space in single pass
		\item {\bf ICB}: Use Iterative Context Bounding\related{Musuvathi '07, Coons '13} to repeatedly test maximal space with increasing preemption bound
			% dont mention BPOR
	\end{itemize}
\end{frame}
